### 2019-08-18 cikk/main/article.tex
added cikk

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tocloft}

\title{SPIE journal papers: sample manuscript showing style and formatting specifications}

\author[a,*]{G\'abor Galg\'oczi}
\author[a]{Second Author}
\author[b]{Third Author}
\author[a,b]{Fourth Author}
\affil[a]{University Name, Faculty Group, Department, Street Address, City, Country, Postal Code}
\affil[b]{Company Name, Street Address, City, Country, Postal Code}

\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table} 
\begin{document} 
\maketitle

\begin{abstract}
Abstract
\end{abstract}

% Include a list of up to six keywords after the abstract
\keywords{Geant4, GRB, gamma-rays, satellite, cosmic background}

% Include email contact information for corresponding author
{\noindent \footnotesize\textbf{*}G\'abor Galg\'oczi,  \linkable{galgoczi.gabor@wigner.mta.hu} }

\begin{spacing}{2}   % use double spacing for rest of manuscript

\section{Introduction}
\label{sect:intro}  % \label{} allows reference to this section

\section{Methodology}

The differential flux is normed. Afterwards the normed integral flux is calculated for each bin. From 0 energy to the given energy.

$$ \int{1}{2} $$

A random number between zero and one is drawn. The bin that has

Check particle energy distirbution

cxb high energy

mizuno raised concern about cxb and gamma albedo??

\section{Simulation of GRB induced signal in the detector}

tbd: what is the background after polar orbits?
casing thickness needs to be optimized for:
more GRB signal
less CXB signal
less electrons...

\section{Results}

\appendix    % this command starts appendixes

\section{Miscellaneous Formatting Details}
\label{sect:misc}

% \disclosures 
\subsection*{Disclosures}

\acknowledgments

%%%%% References %%%%%

\bibliography{report}   % bibliography data in report.bib
\bibliographystyle{spiejour}   % makes bibtex use spiejour.bst

%%%%% Biographies of authors %%%%%

\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering. He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.  He is the author of more than 50 journal papers and has written three book chapters. His current research interests include optical interconnects, holography, and optoelectronic systems. He is a member of SPIE.

\vspace{1ex}
\noindent Biographies and photographs of the other authors are not available.

\listoffigures
\listoftables

\end{spacing}
\end{document}